Tail associations and their impact on extinction risk

### 2019-10-26 MainText.Rmd
elaboration of a comment
moer progress on methods

<!--stochastic matrix model : age-structured-->
To investigate (Q1) from the \nameref{Introduction},
we considered a simple stochastic age-structured model
with two age classes,  
\begin{equation}
\left({\begin{array}{cc} E(t+1) \\ A(t+1) \end{array}}\right) = 
\left(\begin{array}{cc} 0 & f_a(t)\\ s_e(t) & 0 \end{array}\right)
\left(\begin{array}{c} E(t) \\ A(t) \end{array}\right), \label{eq_age_str}
\end{equation}
\noindent where $E(t)$ and $A(t)$ are the numbers or population densities of eggs 
and adults, respectively, at time $t$. We chose a very simple model
because our goal was to carry out an initial evaluation, only, of 
the realism of a potential mechanism. The stochastic vital rates $s_e(t)$ and $f_a(t)$
were selected, independently across times, as follows. 
The egg survival rate $s_e(t)$ was drawn from a beta distribution with parameters
$\alpha=3$ and $\beta=2$. 
Adult fecundity was $f_a(t)=\tilde{f}_a(t)\exp(-A(t)/\Omega)$, where $\tilde{f}_a(t)$
was drawn from a gamma distribution with shape and scale parameters $k=2$
and $\theta=1$, respectively, 
and $\Omega$ is a density dependence parameter. Taking $\Omega=\infty$ made
$\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence
(it was then a standard stochastic matrix model). Finite, positive values of 
$\Omega$ made the model nonlinear, with density dependence in the adult
fecundity rate.

@@

<!--***DAN: Shyamolina, the below text makes me wonder if I am misunderstanding something, please
explain. The part that confuses me is the 2.5th and 97.5th quantile part, what is the meaning
of that and what does it have to do with extinction riks?

@@

<!--density-dependence metapopulation model-->
We explored (Q2) by examining how the nature of the tail association of an
environmental variable across space could influence the extinction risk of a 
metapopulation model with density-dependent feedback. We considered six density-dependent 
population-growth models (Ricker, Hassell, Maynard Smith, Pennycuick, Verhulst, Austin-Brewer)
in noisy environment (see Table \ref{tab_model_summary}) 
from a metapopulation context.
Models were adapted from earlier work
by @cohen1995unexpected, excluding only two: Varley model (this is not a continuous one) 
and the Malthus-Condorcet-Mill model (terms containing logarithm of the extinct population does not make sense). 
All models assumed a generalized form: $\vec{P}_{t+1} = D \lambda_t \vec{P_t}$,
where $\vec{P_t}$ was a vector of population densities, 
the $i^{th}$ component $P_t^i$ representing
the density in the $i^{th}$ habitat patch ($i=1,2,\ldots,n$) at time $t$. 
The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
of stochastic density dependence form $\lambda_i(t)$ (see second column 
of Table \ref{tab_model_summary}). For simplicity, 
we considered $n$ patches in each model were arranged, evenly spaced, in a line
with the provision of "local" or "global" dispersal among patches. An n $\times$ n dispersal
matrix, $D$ assumed that a fraction $d$ of 
each population patch dispersed during any given time step, equally distributed to
the two or one nearest neighbors in case of local dispersal. For global dispersal, the fraction
$d$ dispersed equally to other $(n-1)$ patches. We called $d$ as dispersal
rate with $d$ being zero there was no dispersal, and hence no chance to rescue an extinct 
patch through migration from nearby metapopulation patches.

>>>>

<!--stochastic matrix model : age-structured-->
To investigate Q1 from the \nameref{Introduction},
we considered a simple stochastic age-structured model
with two age classes, 
\begin{equation}
\left({\begin{array}{cc} E(t+1) \\ A(t+1) \end{array}}\right) = 
\left(\begin{array}{cc} 0 & f_a(t)\\ s_e(t) & 0 \end{array}\right)
\left(\begin{array}{c} E(t) \\ A(t) \end{array}\right), \label{eq_age_str}
\end{equation}
\noindent where $E(t)$ and $A(t)$ are the numbers or population densities of eggs 
and adults, respectively, at time $t$. We chose a very simple model
because our goal was to carry out an initial evaluation, only, of 
the realism of a potential mechanism. The stochastic vital rates $s_e(t)$ and $f_a(t)$
were selected, independently across times, as follows. 
The egg survival rate $s_e(t)$ was drawn from a beta distribution with parameters
$\alpha=3$ and $\beta=2$. 
Adult fecundity was $f_a(t)=\tilde{f}_a(t)\exp(-A(t)/\Omega)$, where $\tilde{f}_a(t)$
was drawn from a gamma distribution with shape and scale parameters $k=2$
and $\theta=1$, respectively, 
and $\Omega$ is a density dependence parameter. Taking $\Omega=\infty$ made
$\exp(-A(t)/\Omega)=1$ so that the model did not include density dependence
(it was then a standard stochastic matrix model). Finite, positive values of 
$\Omega$ made the model nonlinear, with density dependence in the adult
fecundity rate.

@@

<!--***DAN: Shyamolina, the below text makes me wonder if I am misunderstanding something, please
explain. The part that confuses me is the 2.5th and 97.5th quantile part, what is the meaning
of that and what does it have to do with extinction riks? I would have thought extinction risk
was just the fraction of 10000 sims that went extinct.

@@

<!--density-dependence metapopulation model-->
To investigate Q2 from the \nameref{Introduction}, we considered spatialized, stochastic versions
of six density dependent population models, the Ricker, Hassell, Maynard-Smith, Pennycuick, Verhulst, 
and Austin-Brewer models as elaborated by @cohen1995unexpected (Table \ref{tab_model_summary}).
All models took the general form $\vec{P} (t+1) = D \lambda(t) \vec{P}(t)$,
where $\vec{P}(t)$ is a vector of population densities, 
the $i^{\text{th}}$ component $P_i(t)$ representing
the population density in the $i^{\text{th}}$ habitat patch ($i=1,2,\ldots,n$) at time $t$. 
The $n \times n$ matrix $\lambda(t)$ was diagonal with $i^{\text{th}}$ diagonal entry 
$\lambda_i(t)$ depnding on the model used (see Table \ref{tab_model_summary}). The matrix $D$
was an $n \times n$ dispersal matrix. For simplicity, 
we assumed the $n$ habitat patches were arranged, evenly spaced, in a line.
We considered both "local" and "global" dispersal among patches in different simulations. 
For local dispersal, a fraction $d$ of 
each population dispersed during each time step, equally distributed to
the two or one nearest neighbors. For global dispersal the same fraction $d$ dispersed
equally to the other $(n-1)$ patches. When $d$ was $0$, dispersal did not occur and 
rescue effects were not possible.
Two models that @cohen1995unexpected used were excluded from our analyses because they were not
readily adapted to our simulation setup: the Varley model,
which is not continuous;
<!--***DAN: Shyamolina, I cannot remember this one and cannot see immediately why non-continuity 
is a reason to exclude these models? I remember we had a good reason to exclude it, I am just 
wanting a bit more explanation here, please. Can you add?-->
and the Malthus-Condorcet-Mill model, which employs a logarithm of the population and therefore
cannot be applied in the context of extinction.
<!--***DAN: go back and look at that model when you get the paper and remind yourself-->

<!--***DAN: got to here 2019 10 25-->
